An antibody-escape estimator for mutations to the SARS-CoV-2 receptor-binding domain

### 2022-03-07 paper/reviewer_response.tex

\response{This is a good question, and we have elaborated on this point at both the beginning of the Results (to explain our approach) and in the Discussion (to mention this as a caveat).
The major reason is that in practice mutation-level measurements tend to be a bit noisier in our data due (for instance) to specific mutations at a site sometimes being at low abundance in the deep mutational scanning libraries.
Averaging over mutations at a site decreases the effects of this noise.
For this reason, we think that the approximation of assuming mutations at a site have similar effects (which is often roughly true) is less bad than the decreased accuracy of the mutation-level compared to site-level measurements.
A secondary reason is that using site-level measurements means that the calculations no longer have to be referred to a specific RBD sequence, which is useful if the immunodominance hierarchy remains the same even as the virus evolves and people are exposed to new variants (for instance, making antibodies against A484 rather than E484 after infection with Omicron).
Overall, the new text should make clear that using the site-level measurements is an approximation for the reason the reviewer points out, but why we think it is overall a useful approximation.}

It is a little difficult to know exactly how the different parameters are being calculated since in many cases they are just describe in the text.  For example, in the Methods section they state that “We also normalize the site-level escape metrics for each antibody to account for difference in strengths of antibody selection” and that “essentially this corresponds to scaling the escape value for each antibody so that a value of one corresponds to the larger of the maximum escape as a site or 20 times the median value across sites”.  It would be much easier to understand how these values are being calculated if the actual equations for each step in the workflow were provided.

\response{This is a good suggestion, and we have added the exact equation used to perform the normalization in the relevant section of the Methods.}

@@

There are several ways in which the binding score could be calculated, e.g., mean versus sum, specific substitutions versus generic substitutions, etc., but the authors only present the results from one approach.  It would be useful to present results from multiple configurations to determine if the results are robust to these configuration parameters.

\response{
This is a good point.
It turns out that using the mean versus sum of mutations at a site gives extremely similar results.
This can be seen graphically by going to the escape calculator at \url{https://jbloomlab.github.io/SARS2_RBD_Ab_escape_maps/escape-calc/} and selecting the bottom metric to be \texttt{mean of mutations at site} rather than \texttt{sum of mutations at site}.
To quantify this more formally, we have also added (as Supplementary Figure 1) a plot of the estimated escape versus experimental measurements using mean at a site in addition to the sum-at-a-site scores shown in Figure 3.
The results are nearly identical: Pearson's R values of 0.91 vs 0.91, 0.72 vs 0.72, 0.88 vs 0.87, 0.82 vs 0.81, 0.76 vs 0.75, and 0.92 vs 0.91.
We have added mention in the text of how the results are similar regardless of whether we use the mean or sum at a site.
}

@@

The most important aspect of the paper is the correlation between the calculated binding score and the experimentally-measure fold change in virus neutralization as presented in Figure 3, which would provide some kind of validation of the computational method.  The authors state in the text that “the escape-calculator scores correlate quite well with the measured neutralization titers across all studies and cohorts”; “quite well” seems pretty subjective.  At a minimum, the strength of correlation should be quantified by using something like a Spearman’s rho metric which could then be used to calculate a t test statistic and p-value.  Indeed, these metrics could be used to compare the performance of the different configuration parameters.

\response{
We have added Pearson's R and the associated P-values in Figure 3.
Five of the 6 correlations are significant, with $P < 0.05$.
In addition, we have used correlation coefficients to compare the sum-of-mutations-at-a-site model to the mean-of-mutations-at-a-site model as described in the point immediately above.
}

### 2022-03-07 paper/paper.tex
add correlations and plot results for site mean as Fig S1

\begin{SCfigure*}[][]
\includegraphics[width=0.7\textwidth]{../results/neut_studies/neut_studies.png}
\caption{Correlation of calculated binding scores with experimentally measured fold changes in neutralization for SARS-CoV-2 variants and mutants (smaller values indicate worse neutralization).
The data of \citet{lucas2021impact} was generated using authentic SARS-CoV-2 and sera from vaccinated individuals who were (top) or were not (bottom) previously infected with SARS-CoV-2.
The data of \citet{uriu2021neutralization} and \citet{wang2021antibody} was generated using pseudovirus against convalescent (top) or vaccine (bottom) sera, with vaccine sera from Pfizer BNT162b2 or Moderna mRNA-1273 vaccines, respectively. 
The fold changes are geometric means over all subjects in each cohort.
An interactive version of this figure that allows mousing over points to see details is at \url{https://jbloomlab.github.io/RBD_escape_calculator_paper/neut_studies.html}.
}
\label{fig:neut_studies}
\end{SCfigure*}

@@

The deep mutational scanning measures an escape fraction for each tolerated RBD mutation against each antibody, which represents an estimate of how completely that mutation escapes antibody binding~\citep{greaney2021complete}.
We summarize the mutation-level escape fractions into site-level measurements in two ways: taking the sum of the mutation escape fractions at each site, or taking the mean of the mutation escape fractions across all tolerated mutations at each site.
The results reported in this paper use the sums as the site-level metric, although the estimator has an option (\texttt{escape\_metric}) to use the mean instead.

@@

\section{Acknowledgments}
We thank collaborators in the Crowe lab, Nussenzweig lab, Bjorkman lab, and at Vir Biotechnology for sharing the antibodies that were used in prior published studies to generate the deep mutational scanning data aggregated here.
We thank the Moyo and de Oliveira labs and other researchers in South Africa and Botswana for rapidly sharing information about the Omicron variant to enable the analyses in the last section of the results.
This work was supported in part by the NIH/NIAID under R01AI141707 (to JDB) and T32AI083203 (to AJG).
The work was also supported by the Gates Foundation grant INV-004949 to JDB.
TNS is a Howard Hughes Medical Institute Fellow of the Damon Runyon Cancer Research Foundation.
JDB is an Investigator of the Howard Hughes Medical Institute.

@@

\bibliography{references}
}
\end{document}

>>>>

\begin{SCfigure*}[][]
\includegraphics[width=0.7\textwidth]{../results/neut_studies/neut_studies.png}
\caption{Correlation of calculated binding scores with experimentally measured fold changes in neutralization for SARS-CoV-2 variants and mutants (smaller values indicate worse neutralization).
The data of \citet{lucas2021impact} was generated using authentic SARS-CoV-2 and sera from vaccinated individuals who were (top) or were not (bottom) previously infected with SARS-CoV-2.
The data of \citet{uriu2021neutralization} and \citet{wang2021antibody} was generated using pseudovirus against convalescent (top) or vaccine (bottom) sera, with vaccine sera from Pfizer BNT162b2 or Moderna mRNA-1273 vaccines, respectively. 
The fold changes are geometric means over all subjects in each cohort.
Labels at the top of each plot shown the Pearson's R and associated P-value.
An interactive version of this figure that allows mousing over points to see details is at \url{https://jbloomlab.github.io/RBD_escape_calculator_paper/neut_studies.html}.
Supplementary Figure S1 shows almost identical results are obtained if the binding scores are computed using the mean of mutations at a site rather than the sum.
}
\label{fig:neut_studies}
\end{SCfigure*}

@@

The deep mutational scanning measures an escape fraction for each tolerated RBD mutation against each antibody, which represents an estimate of how completely that mutation escapes antibody binding~\citep{greaney2021complete}.
We summarize the mutation-level escape fractions into site-level measurements in two ways: taking the sum of the mutation escape fractions at each site, or taking the mean of the mutation escape fractions across all tolerated mutations at each site.
The results reported in this paper use the sums as the site-level metric, although the estimator has an option (\texttt{escape\_metric}) to use the mean instead (the results are almost identical: compare Fig~\ref{fig:neut_studies} versus Supplementary Figure S1.

@@

\section{Acknowledgments}
We thank collaborators in the Crowe lab, Nussenzweig lab, Bjorkman lab, and at Vir Biotechnology for sharing the antibodies that were used in prior published studies to generate the deep mutational scanning data aggregated here.
We thank the Moyo and de Oliveira labs and other researchers in South Africa and Botswana for rapidly sharing information about the Omicron variant to enable the analyses in the last section of the results.
This work was supported in part by the NIH/NIAID under contract HHS75N93021C00015, and grants R01AI141707 (to JDB) and T32AI083203 (to AJG).
The work was also supported by the Gates Foundation grant INV-004949 to JDB.
TNS is a Howard Hughes Medical Institute Fellow of the Damon Runyon Cancer Research Foundation.
JDB is an Investigator of the Howard Hughes Medical Institute.

@@

\bibliography{references}
}

\end{document}
